Place of death in NRS COVID-19 deaths: data processing for analysis, 2023 data

### 2023-02-21 covid_nrs_place_of_death_data_2023.Rmd
modified data to avoid hard-coding current year - lots more to change I'm sure!

```{r}
## a separate raw file makes it easier to debug, and also reduces the number of requests I make while debugging!
ods_raw <- ods_dataset("deaths-involving-coronavirus-covid-19", geography = "sc")

@@

weekly_deaths_cause_of_death_covid <-
  bind_rows(
    weekly_deaths_cause_of_death_2020,
    weekly_deaths_cause_of_death_2021,
    weekly_deaths_cause_of_death_2022_onwards
  ) %>%
  filter(cause_of_death == "COVID-19") %>%
  group_by(year) %>%
  group_map(
    .f = function(data_tbl, year_tbl) {
      current_year <- year_tbl$year  # get year from the function arguments
      year_range <- infer_year_range_for_calculating_covid_deaths(current_year)  # work out which years of data to use
      weekly_means <-  
        weekly_deaths_cause_of_death_covid %>%
        filter(year %in% year_range) %>%  # filter the covid data to the relevant years to compute from
        create_week_53_data_for_years_with_52_weeks() %>%  # create 53rd week if needed
        group_by(place_of_death, cause_of_death, week_number) %>%
        summarise(
          deaths_mean = mean(number_of_deaths, na.rm=TRUE),
          # year_range = paste0(year_range,collapse=","),  # for testing
          .groups = "drop"
        )
      data_tbl %>%
        mutate(year = current_year) %>%  # this gets removed in the grouping, so I re-add it here
        select(-deaths_mean) %>%  # remove the NA/0 mean deaths
        create_week_53_data_for_years_with_52_weeks() %>%  # create 53rd week if needed
        left_join(weekly_means, by = c("place_of_death", "cause_of_death", "week_number"))  # merge the mean deaths
    }
  ) %>%
  bind_rows()  # merge into single tibble from list

@@

# TODO: I get different numbers of 5-year annual average deaths between 2020 and 2021, how come?
historical_annual_average_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  group_by(year, historic_period, place_of_death, cause_of_death) %>%
  summarise(
    annual_deaths_mean = sum(deaths_mean),
    .groups = "drop"
  ) %>% 
  filter(year!=2020)  # 2020 and 2021 should have identical annual means so I remove one %>%
  select(-year)

@@

merged_deaths_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  left_join(  # left_join so that the historical Covid-19 rates are blank!
    historical_average_cause_of_death %>% select(-week_number_run_over), 
    by = c("place_of_death", "week_number", "cause_of_death", "historic_period", "year", "date_w_c")
  )
```

@@

These are used for saying in plots what data were included e.g. weeks 1-12 in 2022.

```{r}
## constants for use in graphs & annotations
weeks_available_2022 <- max(merged_deaths_age$week_number[merged_deaths_age$year==2022])
most_recent_date_available_2022 <- max(merged_deaths_age$date_w_c)

earliest_date <- min(merged_deaths_age$date_w_c)
```

@@

        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2022}, 2022"),

@@

      period = glue::glue("2022 up to week {weeks_available_2022}"),

>>>>

```{r}
## a separate raw file makes it easier to debug, and also reduces the number of requests I make while debugging!
## Note: as of 16 Feb 2023, this returns a strange error
# TODO: if error persists, load data from the main report file instead?
ods_raw <- ods_dataset("deaths-involving-coronavirus-covid-19", geography = "sc")

@@

weekly_deaths_cause_of_death_covid_before_processing <-
  bind_rows(
    weekly_deaths_cause_of_death_2020,
    weekly_deaths_cause_of_death_2021,
    weekly_deaths_cause_of_death_2022_onwards
  )

weekly_deaths_cause_of_death_covid <-
  weekly_deaths_cause_of_death_covid_before_processing %>%
  filter(cause_of_death == "COVID-19") %>%
  group_by(year) %>%
  group_map(
    .f = function(data_tbl, year_tbl) {
      current_year <- year_tbl$year  # get year from the function arguments
      year_range <- infer_year_range_for_calculating_covid_deaths(current_year)  # work out which years of data to use
      weekly_means <-  
        weekly_deaths_cause_of_death_covid_before_processing %>%  # bind again weekly deaths
        filter(year %in% year_range) %>%  # filter the covid data to the relevant years to compute from
        create_week_53_data_for_years_with_52_weeks() %>%  # create 53rd week if needed
        group_by(place_of_death, cause_of_death, week_number) %>%
        summarise(
          deaths_mean = mean(number_of_deaths, na.rm=TRUE),
          # year_range = paste0(year_range,collapse=","),  # for testing
          .groups = "drop"
        )
      data_tbl %>%
        mutate(year = current_year) %>%  # this gets removed in the grouping, so I re-add it here
        select(-deaths_mean) %>%  # remove the NA/0 mean deaths
        create_week_53_data_for_years_with_52_weeks() %>%  # create 53rd week if needed
        left_join(weekly_means, by = c("place_of_death", "cause_of_death", "week_number"))  # merge the mean deaths
    }
  ) %>%
  bind_rows()  # merge into single tibble from list

@@

# TODO: I get different numbers of 5-year annual average deaths between 2020 and 2021, how come?
historical_annual_average_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  group_by(year, historic_period, place_of_death, cause_of_death) %>%
  summarise(
    annual_deaths_mean = sum(deaths_mean),
    .groups = "drop"
  ) %>% 
  filter(year!=2020) %>%  # 2020 and 2021 should have identical annual means so I remove one %>%
  select(-year)

@@

## Note: previously I computed the historical average separately fro mthe weekly cause of death data
## Now it's already computed in the same dataset, so I just assign it here, no joining needed
merged_deaths_cause_of_death <-
  weekly_deaths_cause_of_death
# merged_deaths_cause_of_death <-
#   weekly_deaths_cause_of_death %>%
#   mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
#   left_join(  # left_join so that the historical Covid-19 rates are blank!
#     historical_average_cause_of_death %>% select(-week_number_run_over), 
#     by = c("place_of_death", "week_number", "cause_of_death", "historic_period", "year", "date_w_c")
#   )
```

@@

These are used for saying in plots what data were included e.g. weeks 1-12 in 2023.

```{r}
## constants for use in graphs & annotations
weeks_available_2023 <- max(merged_deaths_age$week_number[merged_deaths_age$year==2023])
most_recent_date_available_2023 <- max(merged_deaths_age$date_w_c)

earliest_date <- min(merged_deaths_age$date_w_c)
current_year <- isoyear(most_recent_date_available_2023)
```

@@

        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2023}, 2022"),

@@

      period = glue::glue("2022 up to week {weeks_available_2023}"),
